---
title: "PA1_template.Rmd"
author: "XINXIN"
date: "8/5/2020"
output: 
    html_document:
        keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r}
library(dplyr)
data <- read.csv("activity.csv")
data <- tbl_df(data)
data$date <- as.Date(data$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r}
stepday <- data %>% group_by(date) %>% summarize(sum = sum(steps))
stepdaywithoutna <- stepday[complete.cases(stepday),]
```
2. Make a histogram of the total number of steps taken each day
```{r}
hist(stepdaywithoutna$sum, main = "Total Number of Steps Taken Each Day", xlab = "Total Number of Steps")
```
3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(stepdaywithoutna$sum)
median(stepdaywithoutna$sum)
```

## What is the average daily activity pattern?

1. Make a time series plot
```{r}
stepavg <- data %>% group_by(interval) %>% summarize(mean = mean(steps, na.rm = TRUE))
plot(stepavg$interval, stepavg$mean, data = stepavg, type = "l", main = "Average Number of Steps Taken", xlab = "Intervals", ylab = "Average Steps")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
stepavg %>% filter(mean == max(stepavg$mean))
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(data$steps))
```
2. Devise a strategy for filling in all of the missing values in the dataset
```{r}
missingdata <- data[is.na(data$steps),]
mer <- merge(missingdata, stepavg)
mer <- arrange(mer, date)
missingdata$steps <- mer$mean
```
3. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
fuldata <- rbind(data[!is.na(data$steps),], missingdata) %>% arrange(date)
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}
stepdayfull <- fuldata %>% group_by(date) %>% summarize(sum = sum(steps))
hist(stepdayfull$sum, main = "Total Number of Steps Taken Each Day", xlab = "Total Number of Steps")
mean(stepdayfull$sum)
median(stepdayfull$sum)
```
There is a little bit different between the mean and median total number of steps taken per day for the data with missing values been filled and the original data.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r}
fuldata$week <- weekdays(fuldata$date)
for (i in 1:nrow(fuldata)){
        if (fuldata$week[i] == "Saturday" ||fuldata$week[i] == "Sunday"){
                fuldata$week[i] = "weekend"
        }else{
                fuldata$week[i] = "weekday"
        }
}
```
2. Make a panel plot containing a time series plot
```{r}
library(ggplot2)
fuldataavg <- fuldata %>% group_by(week, interval) %>% summarize(mean = mean(steps))
qplot(interval, mean, data = fuldataavg, geom = "line", facets = week~.)+labs(title = "The Average Number of Steps taken Across Weekdays and Weekends", y = "Number of Steps")
```